Papers_search

### 2019-01-24 Papers_search.Rmd
after spellcheck

Imagine, you woke up in the morning feeling rather sick. You decided to work from home to get some rest a get better more quickly. You dod not feel energetic enough to crunch numbers or write, so you thought you may catch up on scientific reading instead. You pick up the book sitting on top of the "to-read"" pile and open it at a random place. It happens to be **Ecoimmunology** (edited by Gregory Demas, Randy Nelson, Randy Joe Nelson, 2012) and it opens on the following fragment:

@@

Let's use this hypothetical topicfor the purpos of the exercise. We hope the topic is quite appealing and quiet easy to understand (we all get sick and most of us reproduce, or at least try to).

In this exercise, we will use a few basic tools to illustrate how to perform the first few steps of meta-analysis: refinig the question, searching and screenig literature (these are the same steps as in Systematic Reviews, so the tools are the same and can be useually described as systematic literature/evidence reviewing tools).  
Note that some R packeages (and many other online/software tools) are available (and more always coming) that can help with some of the tasks during this exercise, but there is no single "perfect" tool for the whole process usable for all disciplines and types of questions (althoght some attempts have been made). We simply do not have enough time to introduce more tools and the main purpose of this exercise is to familiarise you with the basic principles/issues that apply when you use any tools. You can always try alternative tools in your own time.

@@

**ANSWER 1:**  
 1. Note that the listed examples involve challenge to immune system - this is becouse experimental manipulation can give clearest answer to whether termnal investment can be happening. So, the more precise  question would be "whether immune challenge can result in terminal investment"
 2. The examples given involve animals (birds), but plants and organisms from other kingdoms can also be used to answer such questions. However, they have quite dfferent immune systems and life histories, so it might be a good idea to restrict the question to animals, or even to a specific animal taxon (birds or insects). Note that in plant research different terminology can be used, e.g. overcompensation (there is a meta-analysis just published in Ecology on overcompensation in response to herbivory: https://www.ncbi.nlm.nih.gov/pubmed/30554427 , and a few older ones).

@@

**ANSWER 2:**  
e) All of the above.   
Doing some scoping searches and some more reading will help to get to know theoretical background, make predictions and get a feel of what sort of empirical papers are available, and order of magnitude of the number of available papers. These searches can reveal mexisting meta-analyses on the topic, key narrative reviews and empirical papers, all of these will help refining the question and deciding on teh inclusion criteria for meta-analysis. We dont have time to do this all during this exercise, so we just do a quick scoping search for existing meta-analyses to familiarise you with GoogleScholar:

**TASK 1:**  
Are there any meta-analyses (or smilar) on terminal investment in animals?   (you can skip the following few tasks if short of time during the exercise)   
In GoogleScholar (https://scholar.google.com) run:  
*"meta-analytic" OR "meta-analysis" "terminal investment"*   
(you can paste this into search box or use Advanced search; see instructions and other GoogleScholar tips here: https://guides.library.ucsc.edu/c.php?g=745384&p=5361954 and https://scholar.google.com/intl/en/scholar/help.html).

You should get around 300 results. By skimming first few pages you can see a few meta-analyses on stress, diet, senescence, parasitism etc., but none probably specifically on terminal investment. There are many empirical papers coming up in this search because they mention meta-analysis in their body text or reference list - note that GoogleScholar searches full text by default. Also by default, the returned results are arranged in order of relevance to the search (whatever that means - its a complex algorithm that decides) and a few first pages contain some potentially relevant empirical papers.

@@

Other things you should know about GoogleScholar - its search interface is rather simplistic and restrictive, it shows only up to 1,000 results for any particular search query, ranking algorithms are obscured, its hard to export the references, the literature coverage may fluctuate (its based on web scraping) and the searches may not be reproducible. Becouse of this it is not really recommended for the main searches in systematic reviews. On the upside, it can be helpful for additional searches and finding grey literature (not published in peer-review journals).

**RESOURCES:**   
Other software useful for exploratory (scoping) ad additional searches (check them up later, also note that they may have similar disadvantages to these of GoogleScholar):
 * SemanticScholar: https://www.semanticscholar.org/
 * Dimensions: https://www.dimensions.ai/
 * CiteHero: https://citehero.com/

**TASK 3:**   
Run another Google Scholar quick search:  
*"terminal investment" reproduction*   
This time you should get around 1900 records. Note that Goofgle Scholar does so called Automated Query Expansion for terms that are not set to be exact phrases (in quotes) - it uses alternative forms of these terms.

**TASK 4:**   
Try Google Scholar quick search:   
*"terminal investment" "immune challenge"*  
This time you should get around 400 records. Skim throug the first three pages of the results to get a feel of the diversity of the studies found (e.g. what taxa are present and what they might measure as reproductive investment) and note the wording of their titles.

@@

Formulate meta-analytic question "whether immune challenge can result in terminal investment" using PICO terms (actually this fits really well becouse we can use experimental data).

**ANSWER 3:**   
  * P = Population                = animal species   
  * I = Intervention              = immune challenge   
  * C = Comparison/Control group  = unchallenged group of animals, otherwiese in the same state   
  * O = Outcome                   = reproduction-related traits   
(further restrictions for all of these can ba added here or later at the criteria stage - we will add them later)

**RESOURCES:**   
An extensive list of alternative systematic review formulation question framweworks here:  https://www.networks.nhs.uk/nhs-networks/nwas-library-and-information-service/documents/alternative-question-structures-for-different-types-of-systematic-review    )

@@

**QUESTION 4:**  
Timespan: should we include studies from any year?

@@

**ANSWER 5:**   
Most meta-analyses only include studies published in English by default. However, if possible it is recommend to include studies also in other languages (usually not many are found, unles you also perform your searches in other languages). 
**Note:** publication statsu can be also used as a criterium - are you happy to include some types of grey literature (e.g. conference proceedings, theses)?

@@

**QUESTION 7:**  
Taxa: should we include any type of taxa?   
**ANSWER 7:**   
Thats a difficult one. Animal kingdom is vast and diverse, and some species are especially odd (e.g. humans). Does it make sense to include single-cell animals? Does it make sense to include humans? Probably not.
How about multicellular animals, except humans? If many studies ascross multiple taxa exist, we could focus on the most frequently studied taxa or biologically most interesting. Alternatively just vertebrates, except humans? Another thing worth being specific about is that the animals should be wild-type (non-mutant, not disease models) and ideally not subject to any other significant manipulations that could affect their response (e.g. drug treatment; note there is a thin fuzzy line what these manipulations are). There might also be reasons to exclude standard laboratory species (mice, rat, yeast, fruitlfy) or maybe focus just on them? But, for now, we will stick to any multicellular animals, except humans.

@@

**QUESTION 9:**   
Control gorup: should we accept any type of control group?  
**ANSWER 9:**   
Control group (unmanipulated) should consist of smimilar individuals, as far as reasonable and not be subject to any other maipulations.

**QUESTION 10:**   
Outcome type: should we accept any reproduction-related measurements and what that means?   
**ANSWER 10:**    
What is reproduction-related trait will depend on the species studied (may require additional reading or relying on the assertions of the authors). Different taxa will have different suites of reproduction related-traits, but these can be grouped into broader categories (e.g. mate attraction includes vocal signals, visual ornaments, pheromone production etc.; reproductive fitness will include number and quality of offspring) and different specific measurments will be available for different taxa and in different studies - this can be quite messy. If we do not have many studies on the topic and very specific question (e.g. "is the number of offspring affected?"") it is ok to be inclusive and later on try to figure out whether some traits are affected more than others.

**QUESTION 11:**  
Data type: should we accept any type of data for meta-analysis?   
**ANSWER 11:**   
It depends on the effect size we are planning to use. In our project it can be Cohens d (Hedges g) or lnRR, which are typically calcualted from mean, SD or SE and N (sample sizes). However, thay can be also inferred from some of the test statistics, or even proportions, if means and other descriptive statistics are not available. Other considerations include what to do if some data is missing or unclear.   
<br>

@@

**QUESTION 13:**  
How many decision trees do we need, and why, for the literature screening process?   
**ANSWER 13:**    
We need two decision trees! (one for each stage).  
This is because abstracts, titles and keywords do not contein enough information to conclusively answer all of the inclusion criteria questions - you should be excluding every paper for which you cannot answer "yes"" to all the questions. Thus this first stage of screening hes to be more tentative, and we would use a more permissive and simpler decision tree. A good tree starts from the questions that are easiest to answer (if we had any year or language restrictions we would put these on the top). Note thet during abstract screening we exclude the records where definete "no" can be inferred from available information, but we allow for some uncertainity for the "yes" answers by allowing "proabably" as an acceptable answer for inclusion, so we can cghheck for more info in the full text.

@@

**ANSWER 14:**  
Most likely this will not be reported in the abstract, or only the measurments that were found to be significant will be mentioned This question needs to be answered during the full-text screening phase.

Full-text screening is more strict and we want to have answers to all our inclusion criteraia. In principle we only have "yes" and "no" answers to the screening questions. However, in practice, if a study looks very relevant but some information is ambiguous or missing, you may decide to contact the authors to get some extra information or data. in such case, you can provisionary include such papers and make final decision once you get, or not, the replies.

@@

Normally, we attempt to do a fairly comprehensive literature search (a full systematic review). The reason is that we want to have as many relevant data points as possible to increase the power of statistical analyses. The other reason is that we make effort to capture obscure studies that potentially are more likely to report non-significant results, and this way we actively try to reducte publication bias in our data set. Search strategy should include using at least two wide-coverage/interdisciplinary databases (we recommend Scopus and Web of Science), and a few additional sources of literature such as forward and backward reference searching (looking at cited and citing papers), publication lists of the key authors or labs, searching for unpublished studies and datasets (we will not cover these here).

In this exercise we only do a bit of the core search using 2 cdatabases:
Scopus and Web of Science.

@@

**TASK 5:**   
Using Scopus **Advanced Search** run this search strings:  
 1. *TITLE-ABS-KEY ( "terminal investment"  AND  "immune challenge" )*    
(approx. 14 hits - search terms too restrictive)   
 2. *TITLE-ABS-KEY ( ( "terminal investment"  OR  "reproductive effort" )  AND  ( "immune challeng\*"  OR  "immunochalleng\*" ) )  *
(approx. 25 hits - a bit better, but still too restrictive)   
 3. *TITLE-ABS-KEY ( ( "terminal investment"  OR  "reproductive effort" )  AND  ( "immune challeng\*"  OR  "immunochalleng\*"  OR  "infect\*" ) )  *
(approx. 180 hits - much better number, but we probably now have included many non-experimental studies while still missing relevant ones)    
 4. *( TITLE-ABS-KEY ( ( "terminal investment"  OR  "reproductive effort" )  AND  ( "immune challeng\*"  OR  "immunochalleng\*"  OR  "infect\*" ) )  AND NOT  TITLE-ABS-KEY ( load  OR  human  OR  people ) )* 
(approx. 144 hits - by adding a group of exclusion tems (AND NOT) we probably got rid of some irrelevant ones).

Ideally, we shoult think and test some more terms for inclusion, after doing some scoping searches, reading reviews and sentinel relevant empirical papers. We stop for now and assume that the string above is our final searxh string. Note that Scopus Advanced Search stores all your searches ans search strings in a neat table at the bottom of the search window - its easy to copy this out and include in the documentation on how you developed your search strategy (same applies to WoS database). The \* symbol stands for wildcart and indicates "any symbols" since these databases do not do Automatic Query Expansion.

**TASK 6:**    
We can easily translate our final search string from Scopus into Web of Science (WoS) **Advanced Search**. Run this search string:    
*TS = (( ( "terminal investment" OR "reproductive effort" ) AND ( "immune challeng\*"  OR  "immunochalleng\*"  OR  "infect\*"  ) ) NOT ( load OR human OR people ) )*  
(approx. 249 hits - more than from Scopus).   WoS stores your search history and also wich sub-data bases were included in the search: Indexes=SCI-EXPANDED, SSCI, A&HCI, CPCI-S, CPCI-SSH, BKCI-S, BKCI-SSH, ESCI, CCR-EXPANDED, IC Timespan=All years" (you can restrit these in Advanced Search)

Both Scopus and WoS allow exporting the found records (with limitations on the max. number per export, **remember to tick the boxes for exporting abstracts!**) - there are multiple file formats you can use, but .ris and .bib are most commonly used ones  - you will need to figure out which ones work for your reference manager software. The next step would be downloading all found records from your "final" search into reference manager of your choice. Combine the references from two databasesinto a single project/folder and removing duplicates (most reference managers have such function, but it never works perfectly - do a visual check after sorting by titles!) before proceeding to screening stage.

@@

Theoretically you could do screening in teh database browser or your reference manager, but since it is recommended that 2 people do the screening independently (or at least part of the screening), it is good to use a dedicated tool to manage this process.

@@

**Note:* it is possible to deduplication algorythms in these are not very good, so its advisable to deduplicate and check your records in the reference manager beforeuploading to the screening software.

@@

We will use these ten referneces to practice title and abstract screening in Rayyan. Screening can be also done in Abstracker, but we found Rayyan bit more relaible and more user-friendly.   
If you are warking in pairs or small groups, one person should create a new project in Rayyan (see below) and invite another person/people to the project using their email address.

 1. Open Rayyan (we assume you already created a free account) and log in.    
 2. My Reviews -> New REVIEW...   
 3. Enter project title, e.g. "Terminal Investment"" and description, e.g. "test"."   
 4. Select files -> upload *Scopus10.ris* and press Continue.     
It may take several minutes for larger files and slower connections, but eventually the uploaded references should appear in the left panel unde "Search methods" (you can upload more files with references via "Add new" link). Click on the "Uploaded References [scopus10.ris]" link in this panel to see the references in the right panel. Generally clicking on the links within panels on the left will filter the records in the right panels to include on higlithed sets (filtering, to get rid of the filter click on it again to remove hihlighting).  
In the left panels there are several sub-panels.  
 5. In the "Keywords for include"" you can add (via "Add new" link) folowing terms (or at least a few of these): challenge, challenged, injected, injection, infected, infection, PHA, SRBC, vaccine (these terms will be higlighted green in the titles and abstracts to help them find more easily; Rayyan tries to populate this list by itself - but its algorythm is medical-centered so you should adjust this list for your projects).   
 6. "Keywords for exclude" - " set to: review, reviewed, model, human, people, women, men (these terms will be higlighted red in the titles and abstracts to help them find more easily; Rayyan tries to populate this list by itself - but its algorythm is medical-centered so you should adjust this list for your projects).    
 7. Click on "All reviews" button in the top right corner to go to the project lists, then open your=project again by chosing it from the list. A view with a pie chart (your screening progress) and a few buttons will appear. If you need to invite other prople to do screening in parallel with you, do that by clicking on the "Invite" button. Stick with "Collaborator" selection and enter the email and message (e.g. "help me!"). Opne your project again - you should see your collaborators added with some info on their work so far. Press on "Show" button to see the references view.  
 8. Click on the first review from the top - an abstract and other record info will appear in the bottom panel.  
 9. Using pour abstract screening decision tree decide whether include this paper for the next stage of screening and press "Include" or "Exclude" button accordingly. And do the assessment for the remaining records. As you go you can create and assign new labels to individual records, e.g.: "unsure/check". You could also enter reasons for exclusion, e.g. "human" (we normally dont do this as there are often hundreds or thousands esxcluded records per project at this stage).  
10. Go to "All reviews" and check if othe rpeople finished their screening, if applicable. After that you can lift the "blind" mode, export all the records with decisions into a .csv files. For collaborative screens rayyan will now provide filters to disply all consensus included, consensus excluded and records with conflicting decisions. If you have any "conflicts" have a look again and discuss why you couldnt agree on a given record , come to agreement and make final decision. Export the records with final consensus decisions, for documentation.

**QUESTION 15:** 
How many papers you excluded? How many conflicts you got if you collaboratively screened (and if so what was cousing these conflicts, can the decision tree be improved?)

**ANSWER 15:**   
We think 4 should have been excluded: Simmons2009 and Clancy 2013 as human studies, Gershman2010 and Gonzales-Tokman 2013 as not likely to have any reprodaction-related measurements.

**TASK 8:**  
Its time to do the full-text screening for the papers that passed totle and abstract screening. There are some full-text papers that passed the intial screening in the directory - assess Nielsen2012 and Ardia2005 using our full-text screening decision tree.

@@

It is very important to keep detailed record of the whole searching, screening and data extraction process. Searching and screening are usually represented as some variant of PRISMA diagram.
With our 144 records from Scopus, 249 from Wos (these two databases are our "main search"), and imaginary around 400 records from other searches we might have done if we had more time, we would proabably end up with something like this one:

@@

**Note** When you exclude records in the full-text screen keep recore for each indicvidual paper on the main reason for exclusion. Later present this information as a supplementary table for your meta-analyses. (Table of included studies should go in the main text, if possible).

Can you think of any other good questions for this section?

>>>>

Imagine, you woke up in the morning feeling rather sick. You decided to work from home to get some rest a get better more quickly. You don't not feel energetic enough to crunch numbers or write, so you thought you may catch up on scientific reading instead. You pick up the book sitting on top of the "to-read"" pile and open it at a random place. It happens to be **Ecoimmunology** (edited by Gregory Demas, Randy Nelson, Randy Joe Nelson, 2012) and it opens on the following fragment:

@@

Let's use this hypothetical topic for the purpose of the exercise. We hope the topic is quite appealing and quiet easy to understand (we all get sick and most of us reproduce, or at least try to).

In this exercise, we will use a few basic tools to illustrate how to perform the first few steps of meta-analysis: refining the question, searching and screening literature (these are the same steps as in Systematic Reviews, so the tools are the same and can be usually described as systematic literature/evidence reviewing tools).  
Note that some R packages (and many other online/software tools) are available (and more always coming) that can help with some of the tasks during this exercise, but there is no single "perfect" tool for the whole process usable for all disciplines and types of questions (although some attempts have been made). We simply do not have enough time to introduce more tools and the main purpose of this exercise is to familiarize you with the basic principles/issues that apply when you use any tools. You can always try alternative tools in your own time.

@@

**ANSWER 1:**  
 1. Note that the listed examples involve challenge to immune system - this is because experimental manipulation can give clearest answer to whether terminal investment can be happening. So, the more precise  question would be "whether immune challenge can result in terminal investment"
 2. The examples given involve animals (birds), but plants and organisms from other kingdoms can also be used to answer such questions. However, they have quite different immune systems and life histories, so it might be a good idea to restrict the question to animals, or even to a specific animal taxon (birds or insects). Note that in plant research different terminology can be used, e.g. overcompensation (there is a meta-analysis just published in Ecology on overcompensation in response to herbivory: https://www.ncbi.nlm.nih.gov/pubmed/30554427 , and a few older ones).

@@

**ANSWER 2:**  
e) All of the above.   
Doing some scoping searches and some more reading will help to get to know theoretical background, make predictions and get a feel of what sort of empirical papers are available, and order of magnitude of the number of available papers. These searches can reveal existing meta-analyses on the topic, key narrative reviews and empirical papers, all of these will help refining the question and deciding on the inclusion criteria for meta-analysis. We don't have time to do this all during this exercise, so we just do a quick scoping search for existing meta-analyses to familiarize you with Google Scholar:

**TASK 1:**  
Are there any meta-analyses (or similar) on terminal investment in animals?   (you can skip the following few tasks if short of time during the exercise)   
In Google Scholar (https://scholar.google.com) run:  
*"meta-analytic" OR "meta-analysis" "terminal investment"*   
(you can paste this into search box or use Advanced search; see instructions and other Google Scholar tips here: https://guides.library.ucsc.edu/c.php?g=745384&p=5361954 and https://scholar.google.com/intl/en/scholar/help.html).

You should get around 300 results. By skimming first few pages you can see a few meta-analyses on stress, diet, senescence, parasitism etc., but none probably specifically on terminal investment. There are many empirical papers coming up in this search because they mention meta-analysis in their body text or reference list - note that Google Scholar searches full text by default. Also by default, the returned results are arranged in order of relevance to the search (whatever that means - its a complex algorithm that decides) and a few first pages contain some potentially relevant empirical papers.

@@

Other things you should know about Google Scholar - its search interface is rather simplistic and restrictive, it shows only up to 1,000 results for any particular search query, ranking algorithms are obscured, its hard to export the references, the literature coverage may fluctuate (its based on web scraping) and the searches may not be reproducible. Because of this it is not really recommended for the main searches in systematic reviews. On the upside, it can be helpful for additional searches and finding grey literature (not published in peer-review journals).

**RESOURCES:**   
Other software useful for exploratory (scoping) ad additional searches (check them up later, also note that they may have similar disadvantages to these of Google Scholar):
 * Semantic Scholar: https://www.semanticscholar.org/
 * Dimensions: https://www.dimensions.ai/
 * CiteHero: https://citehero.com/

**TASK 3:**   
Run another Google Scholar quick search:  
*"terminal investment" reproduction*   
This time you should get around 1900 records. Note that Google Scholar does so called Automated Query Expansion for terms that are not set to be exact phrases (in quotes) - it uses alternative forms of these terms.

**TASK 4:**   
Try Google Scholar quick search:   
*"terminal investment" "immune challenge"*  
This time you should get around 400 records. Skim through the first three pages of the results to get a feel of the diversity of the studies found (e.g. what taxa are present and what they might measure as reproductive investment) and note the wording of their titles.

@@

Formulate meta-analytic question "whether immune challenge can result in terminal investment" using PICO terms (actually this fits really well because we can use experimental data).

**ANSWER 3:**   
  * P = Population                = animal species   
  * I = Intervention              = immune challenge   
  * C = Comparison/Control group  = unchallenged group of animals, otherwise in the same state   
  * O = Outcome                   = reproduction-related traits   
(further restrictions for all of these can be added here or later at the criteria stage - we will add them later)

**RESOURCES:**   
An extensive list of alternative systematic review formulation question frameworks here:  https://www.networks.nhs.uk/nhs-networks/nwas-library-and-information-service/documents/alternative-question-structures-for-different-types-of-systematic-review    )

@@

**QUESTION 4:**  
Time-span: should we include studies from any year?

@@

**ANSWER 5:**   
Most meta-analyses only include studies published in English by default. However, if possible it is recommend to include studies also in other languages (usually not many are found, unless you also perform your searches in other languages). 
**Note:** publication status can be also used as a criterium - are you happy to include some types of grey literature (e.g. conference proceedings, theses)?

@@

**QUESTION 7:**  
Taxa: should we include any type of taxa?   
**ANSWER 7:**   
That is a difficult one. Animal kingdom is vast and diverse, and some species are especially odd (e.g. humans). Does it make sense to include single-cell animals? Does it make sense to include humans? Probably not.
How about multicellular animals, except humans? If many studies across multiple taxa exist, we could focus on the most frequently studied taxa or biologically most interesting. Alternatively just vertebrates, except humans? Another thing worth being specific about is that the animals should be wild-type (non-mutant, not disease models) and ideally not subject to any other significant manipulations that could affect their response (e.g. drug treatment; note there is a thin fuzzy line what these manipulations are). There might also be reasons to exclude standard laboratory species (mice, rat, yeast, fruit fly) or maybe focus just on them? But, for now, we will stick to any multicellular animals, except humans.

@@

**QUESTION 9:**   
Control group: should we accept any type of control group?  
**ANSWER 9:**   
Control group (unmanipulated) should consist of similar individuals, as far as reasonable and not be subject to any other manipulations.

**QUESTION 10:**   
Outcome type: should we accept any reproduction-related measurements and what that means?   
**ANSWER 10:**    
What is reproduction-related trait will depend on the species studied (may require additional reading or relying on the assertions of the authors). Different taxa will have different suites of reproduction related-traits, but these can be grouped into broader categories (e.g. mate attraction includes vocal signals, visual ornaments, pheromone production etc.; reproductive fitness will include number and quality of offspring) and different specific measurements will be available for different taxa and in different studies - this can be quite messy. If we do not have many studies on the topic and very specific question (e.g. "is the number of offspring affected?"") it is OK to be inclusive and later on try to figure out whether some traits are affected more than others.

**QUESTION 11:**  
Data type: should we accept any type of data for meta-analysis?   
**ANSWER 11:**   
It depends on the effect size we are planning to use. In our project it can be Cohen's d (Hedges g) or lnRR, which are typically calculated from mean, SD or SE and N (sample sizes). However, they can be also inferred from some of the test statistics, or even proportions, if means and other descriptive statistics are not available. Other considerations include what to do if some data is missing or unclear.   
<br>

@@

**QUESTION 13:**  
How many decision trees do we need, and why, for the literature screening process?   
**ANSWER 13:**    
We need two decision trees! (one for each stage).  
This is because abstracts, titles and keywords do not contain enough information to conclusively answer all of the inclusion criteria questions - you should be excluding every paper for which you cannot answer "yes"" to all the questions. Thus this first stage of screening hes to be more tentative, and we would use a more permissive and simpler decision tree. A good tree starts from the questions that are easiest to answer (if we had any year or language restrictions we would put these on the top). Note that during abstract screening we exclude the records where definite "no" can be inferred from available information, but we allow for some uncertainty for the "yes" answers by allowing "probably" as an acceptable answer for inclusion, so we can check for more info in the full text.

@@

**ANSWER 14:**  
Most likely this will not be reported in the abstract, or only the measurements that were found to be significant will be mentioned This question needs to be answered during the full-text screening phase.

Full-text screening is more strict and we want to have answers to all our inclusion criteria. In principle we only have "yes" and "no" answers to the screening questions. However, in practice, if a study looks very relevant but some information is ambiguous or missing, you may decide to contact the authors to get some extra information or data. in such case, you can provisionary include such papers and make final decision once you get, or not, the replies.

@@

Normally, we attempt to do a fairly comprehensive literature search (a full systematic review). The reason is that we want to have as many relevant data points as possible to increase the power of statistical analyses. The other reason is that we make effort to capture obscure studies that potentially are more likely to report non-significant results, and this way we actively try to reduce publication bias in our data set. Search strategy should include using at least two wide-coverage/interdisciplinary databases (we recommend Scopus and Web of Science), and a few additional sources of literature such as forward and backward reference searching (looking at cited and citing papers), publication lists of the key authors or labs, searching for unpublished studies and data sets (we will not cover these here).

In this exercise we only do a bit of the core search using 2 databases:
Scopus and Web of Science.

@@

**TASK 5:**   
Using Scopus **Advanced Search** run this search strings:  
 1. *TITLE-ABS-KEY ( "terminal investment"  AND  "immune challenge" )*    
(approx. 14 hits - search terms too restrictive)   
 2. *TITLE-ABS-KEY ( ( "terminal investment"  OR  "reproductive effort" )  AND  ( "immune challeng\*"  OR  "immunochalleng\*" ) )  *
(approx. 25 hits - a bit better, but still too restrictive)   
 3. *TITLE-ABS-KEY ( ( "terminal investment"  OR  "reproductive effort" )  AND  ( "immune challeng\*"  OR  "immunochalleng\*"  OR  "infect\*" ) )  *
(approx. 180 hits - much better number, but we probably now have included many non-experimental studies while still missing relevant ones)    
 4. *( TITLE-ABS-KEY ( ( "terminal investment"  OR  "reproductive effort" )  AND  ( "immune challeng\*"  OR  "immunochalleng\*"  OR  "infect\*" ) )  AND NOT  TITLE-ABS-KEY ( load  OR  human  OR  people ) )* 
(approx. 144 hits - by adding a group of exclusion terms (AND NOT) we probably got rid of some irrelevant ones).

Ideally, we should think and test some more terms for inclusion, after doing some scoping searches, reading reviews and sentinel relevant empirical papers. We stop for now and assume that the string above is our final search string. Note that Scopus Advanced Search stores all your searches ans search strings in a neat table at the bottom of the search window - its easy to copy this out and include in the documentation on how you developed your search strategy (same applies to WoS database). The \* symbol stands for wildcard and indicates "any symbols" since these databases do not do Automatic Query Expansion.

**TASK 6:**    
We can easily translate our final search string from Scopus into Web of Science (WoS) **Advanced Search**. Run this search string:    
*TS = (( ( "terminal investment" OR "reproductive effort" ) AND ( "immune challeng\*"  OR  "immunochalleng\*"  OR  "infect\*"  ) ) NOT ( load OR human OR people ) )*  
(approx. 249 hits - more than from Scopus).   WoS stores your search history and also which sub-data bases were included in the search: Indexes=SCI-EXPANDED, SSCI, A&HCI, CPCI-S, CPCI-SSH, BKCI-S, BKCI-SSH, ESCI, CCR-EXPANDED, IC Timespan=All years" (you can restrict these in Advanced Search)

Both Scopus and WoS allow exporting the found records (with limitations on the max. number per export, **remember to tick the boxes for exporting abstracts!**) - there are multiple file formats you can use, but .ris and .bib are most commonly used ones  - you will need to figure out which ones work for your reference manager software. The next step would be downloading all found records from your "final" search into reference manager of your choice. Combine the references from two databases into a single project/folder and removing duplicates (most reference managers have such function, but it never works perfectly - do a visual check after sorting by titles!) before proceeding to screening stage.

@@

Theoretically you could do screening in the database browser or your reference manager, but since it is recommended that 2 people do the screening independently (or at least part of the screening), it is good to use a dedicated tool to manage this process.

@@

**Note:* it is possible to deduplication algorithms in these are not very good, so its advisable to deduplicate and check your records in the reference manager before uploading to the screening software.

@@

We will use these ten references to practice title and abstract screening in Rayyan. Screening can be also done in Abstracker, but we found Rayyan bit more reliable and more user-friendly.   
If you are working in pairs or small groups, one person should create a new project in Rayyan (see below) and invite another person/people to the project using their email address.

 1. Open Rayyan (we assume you already created a free account) and log in.    
 2. My Reviews -> New REVIEW...   
 3. Enter project title, e.g. "Terminal Investment"" and description, e.g. "test"."   
 4. Select files -> upload *Scopus10.ris* and press Continue.     
It may take several minutes for larger files and slower connections, but eventually the uploaded references should appear in the left panel under "Search methods" (you can upload more files with references via "Add new" link). Click on the "Uploaded References [scopus10.ris]" link in this panel to see the references in the right panel. Generally clicking on the links within panels on the left will filter the records in the right panels to include on highlighted sets (filtering, to get rid of the filter click on it again to remove highlighting).  
In the left panels there are several sub-panels.  
 5. In the "Keywords for include"" you can add (via "Add new" link) following terms (or at least a few of these): challenge, challenged, injected, injection, infected, infection, PHA, SRBC, vaccine (these terms will be highlighted green in the titles and abstracts to help them find more easily; Rayyan tries to populate this list by itself - but its algorithm is medical-centered so you should adjust this list for your projects).   
 6. "Keywords for exclude" - " set to: review, reviewed, model, human, people, women, men (these terms will be highlighted red in the titles and abstracts to help them find more easily; Rayyan tries to populate this list by itself - but its algorithm is medical-centered so you should adjust this list for your projects).    
 7. Click on "All reviews" button in the top right corner to go to the project lists, then open your=project again by choosing it from the list. A view with a pie chart (your screening progress) and a few buttons will appear. If you need to invite other people to do screening in parallel with you, do that by clicking on the "Invite" button. Stick with "Collaborator" selection and enter the email and message (e.g. "help me!"). Open your project again - you should see your collaborators added with some info on their work so far. Press on "Show" button to see the references view.  
 8. Click on the first review from the top - an abstract and other record info will appear in the bottom panel.  
 9. Using pour abstract screening decision tree decide whether include this paper for the next stage of screening and press "Include" or "Exclude" button accordingly. And do the assessment for the remaining records. As you go you can create and assign new labels to individual records, e.g.: "unsure/check". You could also enter reasons for exclusion, e.g. "human" (we normally don't do this as there are often hundreds or thousands excluded records per project at this stage).  
10. Go to "All reviews" and check if other people finished their screening, if applicable. After that you can lift the "blind" mode, export all the records with decisions into a .csv files. For collaborative screens Rayyan will now provide filters to display all consensus included, consensus excluded and records with conflicting decisions. If you have any "conflicts" have a look again and discuss why you couldn't agree on a given record , come to agreement and make final decision. Export the records with final consensus decisions, for documentation.

**QUESTION 15:** 
How many papers you excluded? How many conflicts you got if you collaboratively screened (and if so what was causing these conflicts, can the decision tree be improved?)

**ANSWER 15:**   
We think 4 should have been excluded: Simmons2009 and Clancy 2013 as human studies, Gershman2010 and Gonzales-Tokman 2013 as not likely to have any reproduction-related measurements.

**TASK 8:**  
Its time to do the full-text screening for the papers that passed title and abstract screening. There are some full-text papers that passed the initial screening in the directory - assess Nielsen2012 and Ardia2005 using our full-text screening decision tree.

@@

It is very important to keep detailed record of the whole searching, screening and data extraction process. Searching and screening are usually represented as some variant of PRISMA diagram.
With our 144 records from Scopus, 249 from WoS (these two databases are our "main search"), and imaginary around 400 records from other searches we might have done if we had more time, we would probably end up with something like this one:

@@

**Note:** When you exclude records in the full-text screen keep record for each individual paper on the main reason for exclusion. Later present this information as a supplementary table for your meta-analyses. (Table of included studies should go in the main text, if possible).

*Can you think of any other good questions for this section?   *
